Hector Permafrost

### 2019-06-04 analysis/paper/paper.Rmd
Add more methods information

## Modeling workflow

@@

The second is the "biome" version, which is the same as the "global" version except that the land carbon sink is divided into two "biomes": "non-permafrost" and "permafrost".
Each of these biomes has its own C pools, fluxes, and parameters.
For this configuration, we varied all of the "global" parameters (previous paragraph) for _both_ biomes ($\beta_{permafrost}$, $\beta_{non-permafrost}$, $Q^10_{RH,permafrost}$, $Q^10_{RH,non-permafrost}$, etc.), as well as the distribution of global pools (vegetation, detritus, soil) and initial primary productivity (`npp_flux0`) aross these two biomes.

@@

@@

For all scenarios, we added the emission as a combination of fossil fuel CO2 emissions (`ffi_emissions` column in Hector) and methane emissions (`CH4_emissions` column in Hector).
Where these were explicitly separated, we added them accordingly.
Where they were presented only in terms of C, we did ???

@@

## Initial conditions

For case 1, we use the Hector defaults [@hartin_2015_simple].

For cases 2 and 3, we assumed the same total pool sizes as case 1, but divided them into the permafrost and non-permafrost biomes according to TODO.

@@

## Global Hector

## Permafrost as a biome

>>>>

## Model configurations

@@

The second is the "biome" version, which is the same as the "global" version except that the land carbon sink is divided into two "biomes": "non-permafrost" and "permafrost".
Each of these biomes has its own C pools, fluxes, and parameters.
For this configuration, we varied all of the "global" parameters (previous paragraph) for _both_ biomes ($\beta_{permafrost}$, $\beta_{non-permafrost}$, $Q^10_{RH,permafrost}$, $Q^10_{RH,non-permafrost}$, etc.), as well as the distribution of global pools (vegetation, detritus, soil) and initial primary productivity (`npp_flux0`) across these two biomes and the relative warming factor (compared to the global average) of the permafrost biome.

@@

## Parameter distributions

We drew parameters from loosely informative distributions that span the range of physical plausibility.
The distributions are shown in the following table (unless otherwise specified, we used the same parameter distributions for global and biome cases):

| Parameter                        | Distribution                    | Hector default value |
|----------------------------------|---------------------------------|----------------------|
| $\beta$                          | Uniform(0, 1)                   | 0.36                 |
| Q^10^~RH~                        | Uniform(0, 10)                  | 2.0                  |
| Litter-detritus fraction         | Beta(3.92, 0.08)                | 0.98                 |
| NPP vegetation fraction          | Dirichlet(**0.35**, 0.60, 0.05) | 0.35                 |
| NPP detritus fraction            | Dirichlet(0.35, **0.60**, 0.05) | 0.60                 |
|----------------------------------|---------------------------------|----------------------|
| Permafrost vegetation C fraction | Beta(1, 2)                      | --                   |
| Permafrost soil C fraction       | Beta(1.1, 1.1)                  | --                   |
| Permafrost detritus C fraction   | Beta(1.1, 1.1)                  | --                   |
| Permafrost warming factor        | 1 + 3 * Beta(5, 5)              | --                   |
|----------------------------------|---------------------------------|----------------------|
| Non-permafrost $\alpha$          | TODO                            | --                   |
| Non-permafrost $\phi$            | TODO                            |                      |
| Permafrost $\alpha$              | TODO                            |                      |
| Permafrost $\phi$                | TODO                            |                      |

We used 5000 Hector simulations for the global case and 10,000 times for the biome case (biome + methane case is TODO).

## Analysis of results

We looked at four output variables:
Atmospheric CO~2~ concentration (`Ca`),
total radiative forcing (`Ftot`),
radiative forcing from CO~2~ alone (`FCO2`),
and
global mean atmospheric temperature (`Tgav`).

First, as a metric of overall parametric uncertainty for each case, we calculated the mean, standard deviation, and 90% and 95% quantiles of the time series for each variable for each simulation.
Second, we evaluated the impact of parameters by plotting the values of each variable at 2100 as a function of input parameter value.

Third, we performed a more rigorous sensitivity and uncertainty analysis based on @lebauer_2013_facilitating.
As a metric of parameter uncertainty, we calculated the coefficient of variation (`cv`) for each input parameter as the ratio of its variance to its mean.
We evaluated sensitivity as follows:
First, we fit a multivariate Generalized Additive Model (GAM; `mgcv::gam` function in R) for each variable as a function of all the input parameters.
Then, we calculate the sensitivity ($s$) to each parameter ($x_i$) as the slope (partial derivative) of the resulting GAM at that parameter's median ($\hat{x}$; discretized at +/- 1% of the median), holding all other parameters ($x_{i+1}..x_n$) constant:

$$ s = \frac{d(GAM(x_i | x_{i + 1} .. x_n))}{d(x_i)} = \frac{GAM(1.01 x_i | ...) - GAM(0.99 x_i | ...)}{1.01 x_i - 0.99 x_i}$$

We then define the "elasticity" ($\epsilon$) (or normalized sensitivity) of the model to that parameter as:

$$ \epsilon = s \div \frac{\hat{y}} {\hat{x}} $$

Finally, we calculate the "partial variance" of each parameter (the fractional contribution of each parameter's uncertainty to overall model predictive uncertainty) by passing the distribution of each parameter through the GAM while holding all other parameters constant, calculating the resulting predictive variance for each parameter, and then normalizing these values relative to the sum of all the variances:

$$ Var(x_i) \approx Var[GAM(x_i | x_{i+1} .. x_n)] $$

$$ pvar(x_i) = \frac{Var(x_i)}{\sum_{i}^n {Var(x_i)}} $$

## Global case

## Biome case

# Miscellaneous notes

@@

@@

@@
